Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2015-11-28 rmarkdown/prelims_paper.Rnw
fixed first part of slides

\title{Data exploration, quality control and statistical analyses of
  ChIP-exo experiments\vspace*{\fill}}

@@

We systematically evaluated diverse aspects of ChIP-exo and found the
following characteristics: First, ChIP-exo's background is quite
different than ChIP-Seq's. Second, although often assumed in ChIP-exo
data analysis methods, the ``peak pair'' assumption doesn't hold
locally in real ChIP-exo data. Third, we compared Paired End (PE)
ChIP-Seq with ChIP-exo and found that both protocols are comparable in
resolution and sensitivity for closely located binding event, but as
the distance between binding event increases ChIP-exo shows higher
sensitivity than PE ChIP-Seq. Finally, for given fixed sequencing
depth, ChIP-exo provides higher sensitivity, specificity and spatial
resolution than PE ChIP-Seq.

@@

While the number of produced ChIP-exo data keeps increasing,
characteristics of ChIP-exo data are not fully investigated
yet. First, DNA libraries generated by the ChIP-exo protocol seem to
be less complex than the libraries generated by ChIP-Seq (Mahony et
al., 2015 \cite{exo_review}), i.e. the number of positions to which
the reads can be aligned has been reduced due to the exonuclease
digestion. Second, although the are roughly the same amount of reads
in both strands, locally there may be more reads in one strand than in
the other. To address this challenges, we suggest a collection of
quality control visualizations to understand which of this biases are
present in ChIP-exo data and globally assess the enrichment and
library complexity of a ChIP-exo sample.We gathered ChIP-exo data from
diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
human and FoxA1 factor in mouse (Serandour et al., 2013
\cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
(\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated by
rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
lab).
% While the number of produced ChIP-exo data keeps increasing,
% characteristics of ChIP-exo data and its analysis are not fully
% investigated yet, including issues of sequence biases inherent to
% ChIP-exo data, choice of optimal statistical methods, and
% determination of optimal sequencing depth. However, currently the
% number of available ChIP-exo data is still limited and their
% sequencing depths are still insufficient for such investigation.

Most of current ChIP-Seq QC guidelines (Landt et al., 2012
\cite{encode_qc}) may not be applicable on ChIP-exo, additionally to
our knowledge there are not established quality control pipelines for
ChIP-exo.Previous ChIP-exo analysis used ChIP-Seq samples to compare
the resolution between experiments (\cite{exo1}, \cite{exo2},
\cite{exoillumina}); Carroll et al., 2014 \cite{carroll.qc} studied
the use of the Strand-Cross Correlation (SCC) (Kharchenko et al., 2008
\cite{strandcc}) and showed that by filtering blacklisted regions the
estimation of the SCC is improved. However, this method requires to
know blacklisted regions in advance which may not be available,
additionally using SCC may not be useful since the peaks that are
attained at the read and fragment length are confused in a typical
ChIP-exo SCC curve.  In our pipeline we propose two out-the-shelf
analysis: an enrichment plot and the normalized local SCC coefficient.

@@

\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics{../figs/diagrams/ChIPexo_diagram.png}
\caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
  second step, where an exonuclease enzyme is added and it digest the
  DNA fragments starting from the 5' end until it finds a crosslink
  protein (this Figure is taken from Furey, 2012 \cite{chipbeyond}).}
\label{fig:exo}
\end{wrapfigure}

@@

ChIP-exo experiments first capture millions of DNA fragments (150 -
250 bp in length) that the protein under study interacts with using
random fragmentation of DNA and a protein-specific antibody. Then,
exonuclease is introduced to trim 5' end of each DNA fragment to a
fixed distance from the bound protein. As a result, boundaries around
the protein of interest constructed with 5' ends of reads are located
much closed to bound protein compared to ChIP-Seq. This step is unique
to ChIP-exo and could potentially provide significantly higher spatial
resolution compared to ChIP-Seq. Finally, high throughput sequencing
of a small region (25 to 100 bp) at 5' end of each fragment generates
millions of reads or tags.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
  \caption{An example of a PE ChIP-Seq peak compared to
    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of the
    transcription initiation complex in \emph{E. Coli}. Gene regions
    on the reverse strand are highlighted in light blue and the region
    was centered respect to ChIP-Seq $\sig$ summit. The library
    complexity is lower in ChIP-exo than in ChIP-Seq but the
    resolution is increased, this can be observed at the $\sig$ peak
    that for ChIP-exo was created as the combination of at least three
    ChIP-exo ``sub-peaks''.}
  \label{fig:exo_example}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = .7\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
%  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
    we can see a slight separation into two strong arms, one
    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
    mapped to only one position in the regions for the ChIP-exo
    experiments divided by condition and biological replicate.}
  \label{fig:enrich}
\end{figure}

@@

\begin{figure}[H]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
      varying sequencing depths}Comparison of the number of candidate
    regions (A), predicted events (B), identified targets (C) and
    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
    ChIP-Seq. RegulonDB annotations are considered as a gold
    standard. A gold standard binding events was deemed identified if
    a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity of
    it.}
  \label{fig:design}
\end{figure}

@@

Last summer, we entered to a contest from ENCODE's functional
characterization group to fit a model to classify enhancers. Roughly
speaking, the dataset consisted of a set of regions
$\mathcal{R}_1,\cdots,\mathcal{R}_n$ and a set of labels
$Y_1,\cdots,Y_n \in \{0,1,\}$. Hence, we built a series of features
$X_1,\cdots,X_p$ and fitted a model. The model was trained by using
10-fold CV and we selected the tunned parameters that maximized the
area under the ROC curve. Then, an independent test set was provided
for which we delivered a set of predictions. However, there are in
total another $m >> n$ regions in the experiment, and the cost of
labeling those $m$ regions may be high (in either the actual monetary
cost of the experiment or the time the scientists spends labeling
those regions). Those $n+m$ regions are H3K27ac peaks, hence they were
ranked and the first $n$ were selected randomly from groups with high,
medium and low signal. This raises the following questions:

@@

In Figure \ref{fig:enrich2} it is shown the same relationship as in
Figure \ref{fig:enrich}, but considering only regions formed where the
reads are being allocated in more than 10 (A) and 30 (B) unique
positions respectively. Both plots show that the vertical arms formed
by regions with low $\mbox{ARC}$ is formed by low complexity regions,
that way suggesting that this segment correspond to the background of
a ChIP-exo experiment.

>>>>

\title{Data exploration, quality control and statistical analysis of
  ChIP-exo experiments\vspace*{\fill}}

@@

We systematically evaluated diverse aspects of ChIP-exo and found the
following characteristics: First, ChIP-exo's background is quite
different than ChIP-Seq's. Second, although often assumed in ChIP-exo
data analysis methods, the ``peak pair'' assumption doesn't hold
locally in real ChIP-exo data. Third, we compared Paired End (PE)
ChIP-Seq with ChIP-exo and found that both protocols are comparable in
resolution and sensitivity for closely located binding event, but as
the distance between binding event increases ChIP-exo shows higher
sensitivity than PE ChIP-Seq. Finally, for a given sequencing depth,
ChIP-exo provides higher sensitivity, specificity and spatial
resolution than PE ChIP-Seq.

@@

While the number of produced ChIP-exo data keeps increasing,
characteristics of ChIP-exo data are not fully investigated
yet. First, DNA libraries generated by the ChIP-exo protocol seem to
be less complex than the libraries generated by ChIP-Seq (Mahony et
al., 2015 \cite{exo_review}), i.e. the possible number of positions to
which the reads can be aligned has been reduced due to the exonuclease
digestion, therefore it is . Second, although the are roughly the same amount of reads
in both strands, locally there may be more reads in one strand than in
the other. To address this challenges, we suggest a collection of
quality control visualizations to understand which of this biases are
present in ChIP-exo data and globally assess the enrichment and
library complexity of a ChIP-exo sample. We gathered ChIP-exo data from
diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
human and FoxA1 factor in mouse (Serandour et al., 2013
\cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
(\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated by
rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
lab).
% While the number of produced ChIP-exo data keeps increasing,
% characteristics of ChIP-exo data and its analysis are not fully
% investigated yet, including issues of sequence biases inherent to
% ChIP-exo data, choice of optimal statistical methods, and
% determination of optimal sequencing depth. However, currently the
% number of available ChIP-exo data is still limited and their
% sequencing depths are still insufficient for such investigation.

Most of current ChIP-Seq QC guidelines (Landt et al., 2012
\cite{encode_qc}) may not be applicable on ChIP-exo, additionally to
our knowledge there are not established quality control pipelines for
ChIP-exo. Previous ChIP-exo analysis used ChIP-Seq samples to compare
the resolution between experiments (\cite{exo1}, \cite{exo2},
\cite{exoillumina}); Carroll et al., 2014 \cite{carroll.qc} studied
the use of the Strand-Cross Correlation (SCC) (Kharchenko et al., 2008
\cite{strandcc}) and showed that by filtering blacklisted regions the
estimation of the SCC is improved. However, this method requires to
know blacklisted regions in advance which may not be available,
additionally using SCC may not be useful since the peaks that are
attained at the read and fragment length are confused in a typical
ChIP-exo SCC curve.  In our pipeline we propose two out-the-shelf
analysis: an enrichment plot and the normalized local SCC coefficient.

@@

\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics{../figs/diagrams/ChIPexo_diagram.png}
\caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
  second step, where an exonuclease enzyme is added and it digests the
  DNA fragments starting from the 5' end until it finds a crosslink
  protein (this Figure is taken from Furey, 2012 \cite{chipbeyond}).}
\label{fig:exo}
\end{wrapfigure}

@@

ChIP-exo experiments first capture millions of DNA fragments (150 -
250 bp in length) that the protein under study interacts with using
random fragmentation of DNA and a protein-specific antibody. Then,
exonuclease is introduced to trim 5' end of each DNA fragment to a
fixed distance from the bound protein. As a result, boundaries around
the protein of interest constructed with 5' ends of reads are located
much closer to the bound protein compared to ChIP-Seq. This step is
unique to ChIP-exo and could potentially provide significantly higher
spatial resolution compared to ChIP-Seq. Finally, high throughput
sequencing of a small region (25 to 100 bp) at 5' end of each fragment
generates millions of reads or tags.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
  \caption{An example of a PE ChIP-Seq peak compared to
    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of the
    transcription initiation complex in \emph{E. Coli}. Gene regions
    on the reverse strand are highlighted in light blue and the region
    was centered respect to the ChIP-Seq $\sig$ summits. The library
    complexity is lower in ChIP-exo than in ChIP-Seq but the
    resolution is increased, this can be observed at the $\sig$ peak
    that for ChIP-exo was created as the combination of at least three
    ChIP-exo ``sub-peaks''.}
  \label{fig:exo_example}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = .7\textwidth]{../figs/for_paper/FoxA1_enrichment.pdf}
  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
%  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
    we can see a slight separation into two strong arms, one
    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
    mapped to only one position in the regions for the ChIP-exo
    experiments divided by condition and biological replicate.}
  \label{fig:enrich}
\end{figure}

@@

\begin{figure}[H]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
      varying sequencing depths.} Comparison of the number of candidate
    regions (A), predicted events (B), identified targets (C) and
    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
    ChIP-Seq. RegulonDB annotations are considered as a gold
    standard. A gold standard binding events was deemed identified if
    a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity of
    it.}
  \label{fig:design}
\end{figure}

@@

Last summer, we entered to a contest from ENCODE's functional
characterization group to fit a model to classify enhancers. Roughly
speaking, the dataset consisted of a set of regions
$\mathcal{R}_1,\cdots,\mathcal{R}_n$ and a set of labels
$Y_1,\cdots,Y_n \in \{0,1,\}$. Hence, we built a series of features
$X_1,\cdots,X_p$ and fitted a model. The model was trained by using
10-fold CV and we selected the tuned parameters that maximized the
area under the ROC curve. Then, an independent test set was provided
for which we delivered a set of predictions. However, there are in
total another $m >> n$ regions in the experiment, and the cost of
labeling those $m$ regions may be high (in either the actual monetary
cost of the experiment or the time the scientists spends labeling
those regions). Those $n+m$ regions are H3K27ac peaks, hence they were
ranked and the first $n$ were selected randomly from groups with high,
medium and low signal. This raises the following questions:

@@

In Figure \ref{fig:enrich2} it is shown the same relationship as in
Figure \ref{fig:enrich}, but considering only regions formed where the
reads are being allocated in more than 10 (A) and 30 (B) unique
positions respectively. Both plots show that the vertical arms formed
by regions with low $\mbox{ARC}$ is formed by low complexity regions,
hence it suggests that this arm may correspond to the background of a
ChIP-exo experiment.

### 2015-11-28 rmarkdown/prelims_paper.tex
fixed first part of slides

\begin{figure}[H]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
      varying sequencing depths}Comparison of the number of candidate
    regions (A), predicted events (B), identified targets (C) and
    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
    ChIP-Seq. RegulonDB annotations are considered as a gold
    standard. A gold standard binding events was deemed identified if
    a binding event was estimated at a $\pm$ 15 vicinity of
    it.}
  \label{fig:design}
\end{figure}

>>>>

\begin{figure}[H]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
      varying sequencing depths.} Comparison of the number of candidate
    regions (A), predicted events (B), identified targets (C) and
    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
    ChIP-Seq. RegulonDB annotations are considered as a gold
    standard. A gold standard binding events was deemed identified if
    a binding event was estimated at a $\pm$ 15 vicinity of
    it.}
  \label{fig:design}
\end{figure}

### 2015-11-28 rmarkdown/slides.Rnw
fixed first part of slides

@@

\title[High Resolution Identification of Protein-DNA Binding Events]{High Resolution Identification of Protein-DNA Binding Events
  and Quality Control for ChIP-exo data}

@@

@@

\section{ChIP-exo procedure}

\begin{frame}
\frametitle{ChIP-exo procedure}
  \begin{figure}[H]
    \centering
\includegraphics{../figs/diagrams/ChIPexo_diagram1.png}
\includegraphics{../figs/diagrams/ChIPexo_diagram2.png}\\
\includegraphics{../figs/diagrams/ChIPexo_diagram3.png}
    \caption{ChIP-exo procedure, the diagram is from Furey, 2012 \cite{chipbeyond}}
    \label{fig:exo_diagram}
  \end{figure}
\end{frame}

\section{ChIP-Seq QC measures}
\begin{frame}
  \frametitle{ChIP-Seq QC measures}

\only<1>{
\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
\rowcolors{1}{}{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{tabularx}{\textwidth}{ @{} c|   Y @{}}
  \toprule
  \textbf{QC measure} & \textbf{Definition} \\
  \midrule
  Nr. reads & Self-explanatory. The higher the better...  \\
  PCR bottleneck Coeff. &  Ratio of number of pos. to which EXACTLY one read maps and number  of pos. to which AT LEAST one read maps \\
  Standarized Std. Dev. &  Normalized Std. Deviation of the sequencing coverage \\
  Strand Cross-Corr. &  $  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]$ \\
  Normalized SCC &  Ratio of max value of SCC and min value of SCC$^*$  \\
  \bottomrule
\end{tabularx}
}

\only<1>{
where $n_c^S$ is the coverage for chromosome $c$ and strand $S$. $r$ is
the Pearson correlation and $w_c$ is the proportion of reads in the
experiment for chromosome $c$
}

\only<2>{
\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
\midrule
\rowcolor{RoyalBlue!20}$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
\hline
\rowcolor{Green!20}ER & Human & - & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
\hline
\rowcolor{Green!20}ER & Human & - & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
\hline
\rowcolor{Green!20}ER & Human & - & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
\hline
\rowcolor{Yellow!20}CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
\bottomrule
\end{tabularx}  
}

@@

\section{Comparison of ChIP-exo and ChIP-seq}

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-seq I}  
\begin{figure}[H]
  \centering  
\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\caption{SCC for CTCF factor in HeLa cell line for ChIP-exo and
  SET-ChIP-Seq }
  \label{fig:scc}
\end{figure}

\begin{itemize}
\item There is a \emph{``phantom peak''} at read length.
\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
\item In ChIP-exo, the \emph{``phantom peak''} and the frag. length
  summit are confounded.
\end{itemize}

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq II}

{\small
\begin{itemize}
\item A shows that high dnesity regions are similar between ChIP-Seq
  and ChIP-exo but background regions are not.
\item The peak-pair assumption doesn't hold in ChIP-exo data, some
  regions show strand-imbalance
\end{itemize}
}

% \only<2>{
% \begin{figure}[H]
% \centering
% \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
%   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
% \end{figure}
% }
\end{frame}

\section{The ChIP-exo QC pipeline}

\begin{frame}[plain]
  \frametitle{The ChIP-exo QC pipeline}

\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
\end{figure}

@@

@@

\begin{frame}
\frametitle{Comparison with ChIP-Seq using dPeak}
\only<1>{
  \begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item Sensitivity is the defined as the proportion of identified
    peaks (regulonDoB \cite{regulonDB} is used as gold-stantard)
  \item Resolution is defined as the min. absolute distance of a
    regulonDB annotation to an est. binding location.
  \end{itemize}
}}

@@

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on them.
\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq, being able to label regions as open or
  closed complexes. We want to improve this analysis by using ChIP-exo
  data, and hopefully detecting intermediate step between this two
  states.
\item Find a optimal strategy for labelling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}

@@

>>>>

\usepackage{multirow}

@@

\title[QC and analysis of ChIP-exo]{Data exploration, quality control
  and statistical analysis of ChIP-exo experiments}

@@

\setbeamertemplate{section in toc}[sections numbered]
\setbeamertemplate{subsection in toc}[subsections numbered]

\setbeamertemplate{section in toc}[square]
\setbeamertemplate{subsection in toc}[square]

@@

\section{Background}

\begin{frame}
\frametitle{ChIP-exo procedure}
  \begin{figure}[H]
    \centering
\includegraphics{../figs/diagrams/ChIPexo_diagram1.png}
\includegraphics{../figs/diagrams/ChIPexo_diagram2.png}\\
\includegraphics{../figs/diagrams/ChIPexo_diagram3.png}
    \caption{ChIP-exo procedure, Furey, 2012 \cite{chipbeyond}}
    \label{fig:exo_diagram}
  \end{figure}
\end{frame}

\section{ChIP-exo Data Exploration and Qualiy Control}

\subsection{ChIP-Seq QC measures}

\begin{frame}[t,plain]
  \frametitle{ChIP-Seq QC measures}

  \begin{itemize}
\only<1>{
  \item Number of reads. Self-explanatory, the higher the better
}
  \item<2-> Number of reads.
\only<2>{
\item PCR bottleneck Coeff. Ratio of number of pos. to which
  {\color{RoyalBlue}EXACTLY} one reads maps and number of pos. to
  which {\color{RoyalBlue}AT LEAST} one reads maps }
\only<2>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth]{../figs/for_paper/PBC_example.pdf}
\end{figure}
}
\item<3-> PCR bottleneck Coeff.
\only<3,4>{
\item Strand Cross-Correlation. 
\begin{align}
  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
\nonumber
\end{align}
}
\only<3>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .4\textwidth]{../figs/for_paper/Kharchenko.png}
\caption{SCC explanation. Kharchenko et al., 2008 \cite{strandcc}}
  \end{figure}
}
\only<4>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
\caption{SCC as QC. Landt et al., 2012 \cite{encode_qc}}
\end{figure}
}
\item<5-> Strand Cross-Correlation.
\only<5>{
\item Normalized Strand Cross-Correlation. Ratio between the SCC when
  the shift is the fragment length and min. value of the SCC.
}
\end{itemize}
\end{frame}

\subsection{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}
\begin{frame}[plain]
\frametitle{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}

\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{table}[H]
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
\textbf{IP} & \textbf{Organism}  & \textbf{Condition} & \textbf{Rep.} & \textbf{Nr. reads} & \textbf{PBC} &  \textbf{NSC} & \textbf{Source} \\
\midrule
$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29   & \multirow{4}{\hsize}{Courtesy of Prof. Landick's lab} \\
$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
\hline
FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 21.452 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 60.661 \\
FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 72.312 \\
\hline
ER & Human & - & 1 & 9,289,835 & 0.8082 & 19.843 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
ER & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
\hline
CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 15.977 & From Rhee and Pugh 2011, \cite{exo1} \\
\bottomrule
\end{tabularx}  
\end{table}

{\small
\begin{itemize}
\item For PBC (human and mouse): 0 - 0.5 severe bottlenecking , 0.5 -
  0.8 moderate bottlenecking, 0.8 - 0.9 mild bottlenecking, 0.9 - 1 no
  bottlenecking.
\item For NSC (human and mouse): $ < 1.1$ is relatively low.
\end{itemize}
}
\end{frame}

\subsection{Comparison of ChIP-exo and ChIP-seq}

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq}

{\small
\begin{itemize}
\item A shows that high density regions are similar between ChIP-Seq
  and ChIP-exo but background regions are not.
\item The peak-pair assumption does hold in ChIP-exo data, but not
  locally since some regions show strand-imbalance.
\end{itemize}
}

@@

\begin{frame}[t]
  \frametitle{Comparison of ChIP-exo and ChIP-seq}  
\begin{columns}
\column{.7\textwidth}
\begin{figure}[H]
  \centering  
\includegraphics[width = .9\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\end{figure}

\column{.3\textwidth}
  \begin{minipage}[c][.6\textheight][c]{\linewidth}
{\small
  {\color{RoyalBlue}Figure:} SCC for CTCF factor in HeLa cell line for
    ChIP-exo and SET-ChIP-Seq.\\
    - {\color{ForestGreen} PBC for rep1 is 0.56}
    - {\color{ForestGreen} PBC for rep2 is 0.94}
  }
\end{minipage}
\end{columns}

{\small
\begin{itemize}
\item There is a \emph{``phantom peak''} at read length.
\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
\item In ChIP-exo, the \emph{``phantom peak''} and the fragment length
  summit are confounded.
\end{itemize}
}
\end{frame}

\subsection{Quality Control pipeline}

\begin{frame}[plain,t]

\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
\end{figure}

@@

\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}

@@

\begin{frame}
\frametitle{Comparison with ChIP-Seq using dPeak}
\only<1>{
  \begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item Sensitivity is the defined as the proportion of identified
    peaks (regulonDB \cite{regulonDB} is used as gold-standard)
  \item Resolution is defined as the min. absolute distance of a
    regulonDB annotation to an est. binding location.
  \end{itemize}
}}

@@

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on them.
\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq, being able to label regions as open or
  closed complexes. We want to improve this analysis by using ChIP-exo
  data, and hopefully detecting intermediate step between this two
  states.
\item Find a optimal strategy for labeling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}

@@

% LocalWords:  regulonDoB
